---
title: "RepData Assignment 1"
author: "MT"
date: "11/25/2020"
output: html_document
---
## Exploring the dataset
Unzipping of file, reading and exploring data.
```{r setup, echo=T}
unzip("activity.zip")
act<-read.csv("activity.csv")
summary(act)
names(act)
head(act)
```

Converting act$date to Date format and generating pairs.
```{r}
act$date<- as.Date(act$date)
summary(act)
pairs(act)
```

Summarising data to sum the steps in each day and storing in stepsday.
```{r echo=T, message=FALSE, warning=FALSE}
library(dplyr)
library(magrittr)
stepsday<- act %>% group_by(date) %>% summarise_each(funs(sum)) 
names(stepsday)<- c("date", "total.steps", "intervals")
str(stepsday)
```

## What is mean total number of steps taken a day?
### Total number of steps per day
```{r echo=T, message=FALSE, warning=FALSE}
library(ggplot2)
g<- ggplot(stepsday, aes( y=total.steps, x=date)) + 
    geom_bar(stat = "identity") + 
    ylab("Total Steps") +
    xlab("Date") +
    ggtitle("Total Steps by Date")
  
g
```

### Calculating the mean and median of steps taken per day.
```{r, echo=T}
summary(stepsday)
```

## What is the average daily activity pattern?
### Plotting time series for average daily pattern and searching for maximum steps
```{r}
library(dplyr)
library(magrittr)
library(ggplot2)
stepsint<- act %>% 
    select(steps, interval)%>% 
    group_by(interval) %>% 
    summarise_each(funs(mean(., na.rm=T)))
    str(stepsint)
g<-ggplot(stepsint, aes(interval, steps)) + 
    geom_line(col="red") + 
    ggtitle("Average number of steps per time interval across all days") +
    xlab("Time") + ylab("Steps")
g

stepsint %>% 
    select(steps, interval) %>%
    filter(steps==max(stepsint$steps))
```
## Imputing Missing Values



